Quantifying and controlling biases in dark matter halo concentration estimates

### 2016-07-18 paper/ms.tex
ready to submit

\begin{abstract}
We use bootstrapping to estimate the bias of concentration
estimates on N-body dark matter halos as a function of particle
number. 
We find that algorithms based on the maximum radial velocity
and radial particle binning tend to overestimate the concentration by 
$15\%-20\%$ for halos sampled with $200$ particles and by $7\%$-$10\%$
for halos sampled with $500$ particles. 
To control this bias at low particle numbers we propose a new
algorithm to estimate halo concentrations based on the integrated mass
profile.   
The method uses the full particle information without any binning,
making it reliable in cases when low numerical resolution becomes a
limitation for other methods.
This method reduces the bias to $1\%-3\%$ for halos sampled with
$200$-$500$ particles.
The velocity and density methods have to use halos with least $\sim
4000$ particles in order to keep the biases down to the same low
level. 
We also show that the mass-concentration relationship could be
shallower than expected from numerical experiments once the biases 
of the different concentration measurements are taken into account.
These results show that bootstrapping and the
concentration estimates based on the integrated mass profile are
valuable tools to probe the internal structure of dark matter halos
in numerical simulations.
\end{abstract}

\keywords{
Galaxies: halos --- Galaxies: high-redshift --- Galaxies: statistics
--- Dark Matter --- Methods: numerical 
}

@@

This motivates us to present a third alternative based on fitting the
integrated mass profile.  
This approach has two advantages with respect to the above mentioned methods.  
It does not involve any data binning and does not throw away data points.  
This translates into a robust estimate even at low resolution/particle numbers.  
Furthermore, since the method does not require any binning, there is no need to
tune numerical parameters.   
This method provides a new independent method to estimate the
concentration parameter.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.49\textwidth]{avg_foff_bolshoi.pdf}
  \includegraphics[width=0.48\textwidth]{sigma_foff_bolshoi.pdf}
\end{center}
\vspace{-0.5cm}
\caption{\emph{Left panel}. Average value of the fractional offset ($f_{\rm
    off}$ (Eq. \ref{eq:f_off}) between the concentration in a halo at 
  high resolution and its bootstrapped versions. 
  Thick (thin) lines correspond to massive halos in the Bolshoi (Via
  Lactea) simulation.
  For each case close to $10000$ bootstrapping samples are used.
  The velocity method noticeably overestimates the concentration up to
  a factor of $0.20$, 
  while the new method only underestimates the concentrations by up to
  a factor of $0.04$.
  \emph{Right panel.}
  1$\sigma$ uncertainties on the offset and the concentration
  estimates from the integrated mass method.  
  Lines show the width between the 14 and 86 percentiles
  of the $f_{\rm off}$ distribution at fixed particle number. 
  The lines include the results for the three methods: maximum
  velocity, density and integrated mass.
  Grey symbols correspond to the $1\sigma$ uncertainty on each of
  the halos in the Bolshoi Simulation as a result of the MCMC
  sampling. 
  To allow a fair comparison with $f_{\rm off}$, this uncertainty has
  been normalized by the preferred concentration value for each halo. 
  In both cases (bootstrapping for individual halos and MCMC fit of a
  halo population) the $1\sigma$ is of the same order with a
  decreasing behavior as $1/\sqrt{N_s}$.
  \label{fig:downsampling}}
\end{figure*}

@@

\subsection{Our proposal: estimate from the integrated mass profile}

@@

We divide the enclosed mass mass $M_i$ and the radii $r_i$ by their
virial values to obtain the dimensionless variables $m_i$ and $x_i$.
Once the mass profile is expressed in dimensionless variables the
concentration is the only free parameter.  
We then use an Affine Invariant Markov Chain Monte Carlo implemented
in the python module {\texttt{emcee}} \citep{emcee} to sample the
likelihood function distribution defined by ${\cal L}(c)\propto
\exp(-\chi^2(c)/2)$ where the $\chi^2(c)$ is written as

\begin{equation}
\chi^2(c)= \frac{1}{N}\sum_{i=1}^{N}[\log m_i - \log m(< x_i;c)]^2,
\end{equation}
%
where $m(<x_i;c)$ corresponds to the values in Eq.(\ref{eq:profile})
at $x=x_i$ for a given value of the concentration parameter $c$ and
the $i$ index sums over all the particles in the numerical profile.
From the $\chi^2$ distribution we find the optimal value of the
concentration and its associated uncertainty.

@@

We use two very different simulations to be sure that the results (bias,
uncertainty on the bias) measured using bootstrapping tests are
independent of the kind of parent simulation.

We use data from the Bolshoi cosmological simulation that follows the
non-linear evolution of a dark matter density field sampled with
$2048^3$ particles over a cubic box of $250\ \hMpc$ on a side.  
The cosmological parameters use a Hubble parameter $h=0.73$, a matter density
$\Omega_m=0.3071$ and a normalization of the power spectrum
$\sigma_8=0.82$. 
The data is publicly available at \url{http://www.cosmosim.org/}.  
Details about the structure of the database and the simulation can be
found in \citep{2013AN....334..691R}.

We use a halo sample containing all the halos located in a cubic
sub-volume of $100$ \hMpc\ on a side.  
From this sample we select all the halos at $z=0$ detected with a
Friends-of-Friends (FoF) algorithm with more than 300 particles,
meaning that the masses are in the interval $4\times 10^{10}\leq
M_{\rm FoF}/\hMsun \leq 10^{14}$.  The FoF algorithm used a linking
length of $0.17$ times the mean inter-particle distance. 
This choice translates into an overdensity $\Delta_h\sim 400-700$
dependent on the halo concentration \citep{More2011}.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.49\textwidth]{concentration_bolshoi.pdf}
  \includegraphics[width=0.49\textwidth]{concentration_bolshoi_corrected.pdf}
\end{center}
\vspace{-0.5cm}
\caption{Mass-concentration relationship for the three different
  methods on the Bolshoi data using only relaxed halos.
  The lines correspond to the median concentration values in each mass
  bin.  
  The shaded region presents 10 and 90 per cent spread. 
  The three methods have a similar spread but for clarity we only show
  the spread for the concentrations estimated using the new method.   
  The dotted line corresponds to fits reported by \citep{Prada2012}.
  \label{fig:concentration}. 
  The left panel shows the raw results coming from each algorithm. 
  The right panel introduces a correction on the velocity and
  integrated mass results following the results of the downsampling
  experiments presented in Figure 
  \ref{fig:downsampling}.} 
\end{figure*}

@@

In our bootstrapping experiments we take halos sampled with at least
$10^{5}$ particles and subsample each halo by factors of $2$ up to
$10^{3}$ to measure the concentration every time the halo is
resampled. 
Every time we keep fixed the virial radius and the center found for
the high resolution halo.
We tested that leaving the virial radius and center free in each
bootstrapping iteration has an effect lower than $1\%$ in the
concentration. 
We extract in total 10000 bootstrapping samples for the Bolshoi
and the Via Lactea simulations.
In the case of the Bolshoi simulation we select 14 massive halos and
create 700 subsamples for each halo. 
For the Via Lactea simulation only one halo is subsampled 10000
times.

The average concentration value for the largest number of particles,
$c_{N_{max}}$, provides with a baseline to compare all the other
results.  
We use the following statistic
\begin{equation}
f_{\rm off} = c_{N} / c_{N_{max}} - 1,
\label{eq:f_off}
\end{equation}
to account for the offset between the concentration at a given
downsampled particle number $c_{N}$ and the baseline $c_{N_{max}}$.
For each method we keep its corresponding concentration $c_{N_{max}}$
value at high resolution.

Figure \ref{fig:downsampling} summarizes our results.  
The plot on the left shows the average value of $f_{\rm off}$ as a
function of the particle number.  
As expected, for large enough sample particle numbers,
$N_{s}>4\times 10^3$, the results of the three algorithms coincide
below the $1\%$ level.
For a lower number of particles, the results deviate from the expected
concentration value.  
For $200$ particles the velocity method overestimates by a factor
$14\%$ while the density method overestimates it by $20\%$.
Around the same sampling scale, the new algorithm shows a more stable
behaviour underestimating the concentration only by a factor of
$1\%$-$2\%$

The line on the same panel shows a fit to the data with the following
functional form 
\begin{equation}
f_{\rm off} = \frac{A}{(1+\log_{10}N_s)^{B}}.
\label{eq:f_off_fit}
\end{equation} 
for the dataset shown in the panel we have $A=2842\pm 1900$,
$B=7.96\pm 0.54$ for the density method; $A=239\pm 131$,
$B=6.23\pm 0.43$ for the velocity method; $A=-3.66\pm 5.22$,
$B=4.65\pm 1.12$ for the integrated mass method.
The uncertainties are estimated from the covariance  matrix resulting
from the non-linear optimization procedure. The lines on Figure
\ref{fig:downsampling} are plotted using these best values.

The right panel in \ref{fig:downsampling} shows different uncertainty
results. 
The lines show the difference between the 14 and 86 percentiles in the
$f_{\rm off}$ distribution at fixed mass.
Each line corresponds to the three different methods to estimate the
concentration and the two simulations.
This shows that the bootstrapping technique can help us to assign a
$1\sigma$ uncertainty to the concentration values at a fixed $N_s$. 
This uncertainty is shown as normalized by the preferred value of the
concentration, which in this case corresponds to $c_{N_{max}}$.  
The circles in the same figure show the $1\sigma$ uncertainty as
estimated from the MCMC chains using the integrated mass method on all
the relaxed halos in the Bolshoi simulation sample. 
To allow for a fair comparison with the bootstrapping results, this
uncertainty was normalized by the preferred concentration.
The MCMC uncertainties follow the same trend as the bootstrapping
results.

From these results we defined a universal $1\sigma$ fractional
uncertainty for the concentration as a function of the number of
particles $N_s$ as  
%
\begin{equation}
\sigma_{c} = \frac{0.40}{\sqrt{N_s/200}}.
\end{equation}
%
An interesting consequence of this result is that $\sigma_c=1$ for
$N_s=32$ particles, independent of the method used to estimate the
concentration.

@@

We now compare the results for the mass-concentration
relationship obtained with the three different algorithms.
This helps to compare the concentration results
of the new algorithm and, more importantly, showcase a probable
consequence of the biases detected through bootstrapping.

Figure \ref{fig:concentration} shows the mass-concentration
relationship for the density, velocity and integrated mass method.
The left panel shows the results as they are produced by each of the
algorithms, the thin dashed line marks the results by
\citep{Prada2012}.

We notice first that the results from the new algorithm follow very
closely the velocity algorithm at high masses ($M_h>10^{12}\hMsun$ or
equivalently for $>8\times10^3$ particles). 
For lower masses there is a difference between the median of the two
methods, but they are still consistent within the statistical
uncertainties.

We also notice that the results from the density method have a
systematic $15\%$ offset from the velocity methods.  
This offset was already presented by \cite{Prada2012} for low
concentrations ($c<6$) and high ($M_h>10^{12}\hMsun$) halo masses.  
Recently \citep{Klypin2016} summarized results for the
mass-concentration relationship coming from different methods and
datasets to show that similar systematic offsets are present.
\citep{2014MNRAS.441.3359D} studied the mass-concentration
relationship using the maximum velocity and density methods and did
not report any significant difference. 
However, they implemented a modified version of the velocity algorithm
that uses the data binning obtained for the density algorithm.

We hypothesize that the increase in the results for the velocity
and density methods below $8\times 10^{3}$ particles come from the
systematic bias described in the previous section.  
To test the general consistency of this hypothesis we correct the
concentration values in the velocity and integrated mass methods by a
factor of $1/(1+f_{\rm  off})$, using the definition in Equation
(\ref{eq:f_off}) and the  
parameters obtained from the data presented in Figure
(\ref{fig:downsampling}).  
The correction brings into perfect agreement the results between the
velocity/density methods and the new algorithm.

@@

These results motivated us to introduce a method with a robust
performance at low particle numbers.
This method is based on the integrated mass profile. 
The results of the algorithm on the same data sample showed a bias  of
$2\%-3\%$ for halos with $200$ particles and less than $1\%$ for halos
with $500$ particles or more.  
To keep the bias of the velocity and density methods below $2\%$ only halos
with at least $\sim 4000$ particles should be taken into account.

@@

 The authors acknowledge the technical support from the new
 high-performance computing facility at Uniandes. JEF-R acknowledges
 financial support from Vicerrector\'ia de Investigaciones at Uniandes
 through a FAPA project. JCMC acknowledges financial support from
 ``Estrategia de  sostenibilidad 2014-2015, Universidad de
 Antioquia''.

>>>>

\begin{abstract}
We use bootstrapping to estimate the bias of concentration
estimates on N-body dark matter halos as a function of particle
number. 
We find that algorithms based on the maximum radial velocity
and radial particle binning tend to overestimate the concentration by 
$15\%-20\%$ for halos sampled with $200$ particles and by $7\%$-$10\%$
for halos sampled with $500$ particles. 
To control this bias at low particle numbers we propose a new
algorithm that estimates halo concentrations based on the integrated
mass profile.
The method uses the full particle information without any binning,
making it reliable in cases when low numerical resolution becomes a
limitation for other methods.
This method reduces the bias to $< 3\%$ for halos sampled with
$200$-$500$ particles.
The velocity and density methods have to use halos with least $\sim
4000$ particles in order to keep the biases down to the same low
level. 
We also show that the mass-concentration relationship could be
shallower than expected once the biases  of the different
concentration measurements are taken into account. 
These results show that bootstrapping and the concentration estimates
based on the integrated mass profile are valuable tools to probe the
internal structure of dark matter halos in numerical simulations. 
\end{abstract}

\keywords{
Galaxies: halos --- Dark matter --- Methods: numerical 
}

@@

This motivates us to present a third alternative based on fitting the
integrated mass profile.  
This approach has two advantages with respect to the above mentioned methods.  
It does not involve any data binning and does not throw away data points.  
This translates into a robust estimate even at low resolution/particle numbers.  
Furthermore, since the method does not require any binning, there is no need to
tune numerical parameters.   
This is a new independent method to estimate the
concentration parameter.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.49\textwidth]{avg_foff_bolshoi.pdf}
  \includegraphics[width=0.48\textwidth]{sigma_foff_bolshoi.pdf}
\end{center}
\vspace{-0.5cm}
\caption{\emph{Left panel}. Average value of the fractional offset ($f_{\rm
    off}$ (Eq. \ref{eq:f_off}) between the concentration in a halo at 
  high resolution and its bootstrapped versions. 
  Thick (thin) lines correspond to massive halos in the Bolshoi (Via
  Lactea) simulation.
  For each case close to $10000$ bootstrapping samples are used.
  The density method noticeably overestimates the concentration up to
  a factor of $20\%$, 
  while the new method only underestimates the concentrations by less
  than $3\%$.
  \emph{Right panel.}
  1$\sigma$ uncertainties on the offset and the concentration
  estimates from the integrated mass method.  
  Lines show the width between the 14 and 86 percentiles
  of the $f_{\rm off}$ distribution at fixed particle number. 
  The lines include the results for the three methods: maximum
  velocity, density and integrated mass.
  Grey symbols correspond to the $1\sigma$ uncertainty on each of
  the halos in the Bolshoi Simulation as a result of the MCMC
  sampling. 
  To allow a fair comparison with $f_{\rm off}$, this uncertainty has
  been normalized by the preferred concentration value for each halo. 
  The $1\sigma$ bootstrapping uncertainty follow a $1/\sqrt{N_s}$
  behaviour. 
  \label{fig:downsampling}}
\end{figure*}

@@

\subsection{Estimate from the integrated mass profile}

@@

We divide the enclosed mass $M_i$ and the radii $r_i$ by their
virial values to obtain the dimensionless variables $m_i$ and $x_i$.
Once the mass profile is expressed in dimensionless variables the
concentration is the only free parameter.

Using bootstrapping data (\S \ref{sec:bootstrapping}) we find that at
a given normalized radius, $x$, the logarithm of the normalized
integrated mass, $m$, approximately follows a gaussian distribution
with variance

\begin{equation}
\sigma_x^2 = \frac{1-x}{x}\frac{1}{N}.
\label{eq:sigma}
\end{equation}

If the integrated mass values at different radii were independent from
each other we could write a likelihood distribution as ${\cal
  L}(c)\propto \exp(-\chi^2(c)/2)$ with

\begin{equation}
\chi^2(c)= \sum_{i=2}^{N-1}\frac{[\log m_i - \log m(< x_i;c)]^2}{\sigma_i^2},
\end{equation}
%
where $\sigma_i^2=\sigma_x^2(x_i)$, $m(<x_i;c)$ corresponds to the
values in Eq.(\ref{eq:profile}) at $x=x_i$ for a given value of the
concentration parameter $c$, and the $i$ index sums over all the
particles in the numerical profile. 
In this computation the particles $i=1$ and $i=N$ are discarded to
avoid divergent terms in the sum.

However, tests on the bootstrapping data show that using
$\sigma_i^2=\sigma_x^2(x_i)$, instead of the full inverse covariance
matrix, grossly overestimates $\chi^2(c)$,
providing small uncertainties around the best concentration value. 
To avoid the expensive computation and inversion of a full covariance
matrix we use the bootsrapping data to calibrate an effective
$\sigma_i^2\approx \sigma_{\rm eff}^2(x_i)$.

We impose two conditions on the approximate $\sigma_{\rm eff}^{2}$. It
must keep the dependence on $x$ that we have discovered for the
diagonal elements and must give similar curves of $\chi^{2}(c)$-$c$
around the minimum as the full covariance matrix.
We found that this effective $\sigma_{\rm eff}^2$ can be approximated
as

\begin{equation}
\sigma_{\rm eff}^2 = \frac{1-x}{x}\frac{N^{1.15}}{4.5\times 10^3}.
\end{equation}

We then use an Affine Invariant Markov Chain Monte Carlo (MCMC)
implemented in the python module {\texttt{emcee}} \citep{emcee} to sample the
likelihood function distribution.
From the $\chi^2$ distribution we find the optimal value of the
concentration and its associated uncertainty. Different choices for
$\sigma^2_{\rm eff}$ do not affect the optimal concentration value,
only its uncertainty.

@@

We use two very different simulations to test our methods.
The first data sample comes from the Bolshoi run, a cosmological
simulation that follows the non-linear evolution of a dark matter density field
sampled with $2048^3$ particles over a cubic box of $250\ \hMpc$ on a side.  
The cosmological parameters use a Hubble parameter $h=0.73$, a matter density
$\Omega_m=0.3071$ and a normalization of the power spectrum
$\sigma_8=0.82$. 
The data is publicly available at \url{http://www.cosmosim.org/}.  
Details about the structure of the database and the simulation can be
found in \citep{2013AN....334..691R}.

We use a halo sample containing all the halos located in a cubic
sub-volume of $100$ \hMpc\ on a side containing a total of $64531$
halos. 
From this sample we select all the halos at $z=0$ detected with a
Friends-of-Friends (FoF) algorithm with more than 300 particles,
meaning that the masses are in the interval $4\times 10^{10}\leq
M_{\rm FoF}/\hMsun \leq 10^{14}$.  The FoF algorithm used a linking
length of $0.17$ times the mean inter-particle distance.  
This choice translates into an overdensity $\Delta_h\sim 400-700$
dependent on the halo concentration \citep{More2011}.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.49\textwidth]{concentration_bolshoi.pdf}
  \includegraphics[width=0.49\textwidth]{concentration_bolshoi_corrected.pdf}
\end{center}
\vspace{-0.5cm}
\caption{Mass-concentration relationship for the three different
  methods on the Bolshoi data using only relaxed halos.
  The lines correspond to the median concentration values in each mass
  bin.  
  The shaded region presents 10 and 90 per cent spread. 
  The three methods have a similar spread but for clarity we only show
  the spread for the concentrations estimated using the new method.   
  The dotted line corresponds to fits reported by \citep{Prada2012}.
  The left panel shows the raw results coming from each algorithm. 
  The right panel introduces a correction on the velocity and
  integrated mass results following the results of the downsampling
  experiments presented in Figure 
  \ref{fig:downsampling}.
  \label{fig:concentration}} 
\end{figure*}

@@

In our bootstrapping experiments we take halos sampled with at least
$10^{5}$ particles and subsample each halo by factors of $2$ up to
$10^{3}$ to measure the concentration every time the halo is
resampled. 
Every time we keep fixed the virial radius and the center found for
the high resolution halo.
We tested that leaving the virial radius and center free in each
bootstrapping iteration has an effect smaller than $1\%$ in the
concentration. 
We extract in total 10000 bootstrapping samples for the Bolshoi
and the Via Lactea simulations.
In the Bolshoi simulation we select 14 massive halos and create 700
subsamples for each one.  
For the Via Lactea simulation the same halo is subsampled 10000
times.

The average concentration value for the largest number of particles,
$c_{N_{max}}$, provides a baseline to compare all the other results.  
We use the following statistic
\begin{equation}
f_{\rm off} = c_{N} / c_{N_{max}} - 1,
\label{eq:f_off}
\end{equation}
to account for the offset between the concentration at a given
downsampled particle number $c_{N}$ and the baseline $c_{N_{max}}$.

Figure \ref{fig:downsampling} summarizes our results.  
The plot on the left shows the average value of $f_{\rm off}$ as a
function of particle number.  
As expected, for a large enough particle number, $N_{s}>4\times 10^3$,
the results of the three algorithms coincide below the $1\%$ level.
For a lower number of particles the results start to deviate.
At $200$ particles the velocity method overestimates the concentration
by a factor $14\%$ while the density method overestimates it by $20\%$.
Around the same sampling scale, the new algorithm shows a more stable
behaviour underestimating the concentration only by a factor of
$1\%$-$3\%$

The thin continous lines on the same panel shows a fit to the data
with the following functional form 
\begin{equation}
f_{\rm off} = \frac{A}{(1+\log_{10}N_s)^{B}}.
\label{eq:f_off_fit}
\end{equation} 
for the dataset shown in the panel we find $A=2842\pm 1900$,
$B=7.96\pm 0.54$ for the density method; $A=239\pm 131$,
$B=6.23\pm 0.43$ for the velocity method; $A=-3.66\pm 5.22$,
$B=4.65\pm 1.12$ for the integrated mass method.
The uncertainties are estimated from the covariance  matrix resulting
from the non-linear optimization procedure. The lines on Figure
\ref{fig:downsampling} are plotted using these best values.

The right panel in Figure \ref{fig:downsampling} shows different uncertainty
results.  
The lines show the difference between the 14 and 86 percentiles in the
$f_{\rm off}$ distribution at fixed mass. 
Each line corresponds to the three different methods to estimate the
concentration and the two simulations.
This shows that the bootstrapping technique can help us to assign a
$1\sigma$ uncertainty to the concentration values at a fixed $N_s$as 
%
\begin{equation}
\sigma_{c} = \frac{0.40}{\sqrt{N_s/200}}.
\end{equation}
%
An interesting consequence of this result is that $\sigma_c=1$ for
$N_s=32$ particles, independent of the method used to estimate the
concentration.

The circles in the same Figure show the $1\sigma$ uncertainty as
estimated from the MCMC chains using the integrated mass method on all
the relaxed halos in the Bolshoi simulation sample. 
To allow for a fair comparison with the bootstrapping results, this
uncertainty was normalized by the concentration value.
Below the range $N_{s}<10^5$ the MCMC uncertainty is lower than the
bootstrapping estimate, with a weak dependence on the particle
number.

@@

We now compare the results for the mass-concentration
relationship obtained with the three different algorithms.
This helps us to study the concentration results
for the new algorithm and, more importantly, showcase a probable
consequence of the biases detected through bootstrapping.

Figure \ref{fig:concentration} shows the mass-concentration
relationship for the density, velocity and integrated mass method.
The left panel shows the results as they are produced by each of the
algorithms. The thin dashed line marks the trend reported by
\citep{Prada2012} using the velocity method.

We notice first that the results from the new algorithm follow very
closely the velocity algorithm at high masses ($M_h>10^{12}\hMsun$ or
equivalently for $>4\times10^3$ particles). 
For lower masses there is a difference between the median of the two
methods, but they are still consistent within the statistical
uncertainties.

We also notice that the results from the density method have a
systematic $15\%$ offset from the velocity methods.  
This offset was already presented by \cite{Prada2012} for low
concentrations ($c<6$) and high ($M_h>10^{12}\hMsun$) halo masses.  
Recently \citep{Klypin2016} summarized results for the
mass-concentration relationship coming from different methods and
datasets to show that similar systematic offsets are present.
\citep{2014MNRAS.441.3359D} studied the mass-concentration
relationship using the maximum velocity and density methods and did
not report any significant difference. 
However, they implemented a modified version of the velocity algorithm
that bins the particle data, which might explain why they the offset
was not reported.

We hypothesize that the increase in the results for the velocity
and density methods below $4\times 10^{3}$ particles comes from the
systematic bias described in the previous section.  
To test the general consistency of this hypothesis, we correct the
concentration values in the velocity and integrated mass methods by a
factor of $1/(1+f_{\rm  off})$, using the definition in Equation
(\ref{eq:f_off}) and the  
parameters obtained from the data presented in Figure
(\ref{fig:downsampling}).  
The correction brings into good agreement the results between the
velocity/density methods and the new algorithm.

@@

These results motivated us to introduce a method with a robust
performance at low particle numbers.
This method is based on the integrated mass profile. 
The results of the algorithm on the same data sample showed a bias  of
$< 3\%$ for halos with $200$ particles and less than $1\%$ for halos
with $500$ particles or more.  
To keep the bias of the velocity and density methods below $2\%$ only halos
with at least $\sim 4000$ particles should be taken into account.

@@

 The authors acknowledge the technical support from the new
 high-performance computing facility at Uniandes. JEFR acknowledges
 financial support from Vicerrector\'ia de Investigaciones at Uniandes
 through a FAPA project. JCMC acknowledges financial support from
 ``Estrategia de  sostenibilidad 2014-2015, Universidad de
 Antioquia''.
